Collective excitations in competing phases in two and three dimensions

### 2023-12-13 paper/main.tex
sdahfousgfousa

In this section, we will be investigating four different diagonal Green's functions $G_{AA^\dagger}(\omega + i0^+)$, namely with
\begin{subequations}
    \label{eqn:resolvent_bases}
    \begin{align}
        A_\text{Higgs} &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( f_{\vk} + f_{\vk}^\dagger \right) \,,\\
        A_\text{Phase} &= \frac{i}{\sqrt{N}} \sum_{\vk} \left( f_{\vk} - f_{\vk}^\dagger \right) \,,\\
        A_\text{CDW}   &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( g_{\vk \up} + g_{\vk \down} \right) \,,\\
        A_\text{AFM}   &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( g_{\vk \up} - g_{\vk \down} \right) \,,
    \end{align}
\end{subequations}
where $N$ is the number of lattice sites in the system. 
Each one of these operators generates a different kind of collective mode.
The first operator captures the amplitude mode of the $s$-wave superconducting state, while the second one catches the phase mode \cite{Fan22}.
The remaining two capture the collective behaviour of the CDW and AFM states, respectively.
Note, that this choice of normalization cancels out with other system size-dependent terms so that our final results do not depend on it at all.

Note additionally, that all of the aforementioned operators are Hermitian. 
This that all spectral functions need to be completely antisymmetrical, see \autoref{sec:antisymmetry_spectral}.
For this reason, we will only show the $\omega \geq 0$ in our plots. 
Additionally, we will add a small imaginary part of $10^{-5}$ to $\omega$ in order to visualize the occurring $\delta$ peaks.

Let us first show and discuss the general form of the spectral functions in the SC and CDW phase and afterward, 
we will describe the behaviour of the occurring modes as the system approaches a phase transition.

The spectral functions corresponding to the individual operators in \eqref{eqn:resolvent_bases} are shown in \autoref{fig:resolvent_overview_SC}.
We chose $U=-2.5$ and $V=\pm0.1$ for both lattices.

@@

\textcolor{red}{Die Festellung, dass die Peaks deltas bzw Ableitungen von deltas sind, könnte ich mir auch in einem Anhang vorstellen.}
To further classify the aforementioned peak that lies outside the two-particle continuum,
we turn towards the real part of the Green's function.
We plot it in close proximity to the peaks with a double logarithmic scale and fit the result linearly to obtain its power-law behaviour.
In practice, we use
\begin{align}
    \ln(\Re[\mathcal{G}](\omega - \omega_0 )) &= a \ln(\omega - \omega_0) + b \nonumber \\ 
    \Leftrightarrow \Re[\mathcal{G}] (\omega) &= (\omega - \omega_0)^a e^b\,,
\end{align}
where $\omega_0$ is the peak's position and $a, b$ are fit parameters.
The real part is related to the imaginary part via the Kramers-Kronig relations. \textcolor{red}{Zitiert man die oder nimmt man an, dass jeder weiß, worum es da geht?}
Specifically, if the imaginary part is a $\delta$ distribution, the real part will be proportional to $1/\omega$, i.e., $a=-1$ and the peak has the weight $\exp(b)$.
Furthermore, an exponent of $a=-2$ corresponds to the derivative of the $\delta$ distribution, i.e., $\delta'(\omega - \omega_0)$.

The plots for the phase peak and the CDW peak in the SC phase can be seen in \autoref{fig:peaks_in_sc}.
The SC peak in the CDW phase behaves similarly to the latter and is therefore not shown explicitly.
We get a good agreement with the power-law behaviour near the peaks with the exponents $a=-2$ (phase) and $a=-1$ (CDW), respectively.
Therefore, we conclude that the CDW peak corresponds to a $\delta$ distribution, while the phase peak is its derivative.
Note, that the latter is located at $\omega=0$ of a perfectly asymmetrical bosonic spectral function and therefore must have 0 weight.
This is given here, because $\int \delta'(\omega) \mathrm{d}\omega = 0$.
The peak itself can be interpreted as two $\delta$ distributions at $0^\pm$.

Next, we will investigate the Higgs peak. 
Additionally, we see in \autoref{fig:resolvent_overview_SC} that the Higgs and phase spectral functions become indistinguishable towards the upper half of the two-particle continuum.
Therefore, we interpret the latter as a background to the Higgs peak itself and compute the difference between both spectral functions.
We plot the result double logarithmically and fit it again linearly, as seen in \autoref{fig:higgs_peak}.
This time around, the exponent is $a\approx-0.5$, indicating that the peak behaves as $1 / \sqrt{\omega - \omega_0}$.

@@

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/resolvent_overview_AFM_CDW.pdf}
    \caption{Spectral functions with respect to the operators listed in \eqref{eqn:resolvent_bases}.
    The left column shows the results for the square lattice while the right column shows the results for the simple cubic lattice.
    %The left edge of the plots is at $-0.05$ rather than $0$ in order to improve the visibility of the phase peak.
    The shaded area marks the two-particle continuum.
    The panels (a) and (b) show the spectral functions in the CDW and AFM phases, respectively.}
    \label{fig:resolvent_overview_AFM}
\end{figure*}

Similar to the previous section, we will first show the general form of the spectral functions with respect to the operators \eqref{eqn:resolvent_bases}.
They are depicted in \autoref{fig:resolvent_overview_AFM}, choosing $V=1$ for the square lattice and $V=0.75$ for the simple cubic lattice.
For the CDW phase (panel a) we chose $U=zV + 0.1$ and for the AFM phase (panel b) $V=zV - 0.1$.

@@

We show that $\mathcal{M}$ is positive semidefinite if the system is in thermal equilibrium.
For any $\vec{x} \in \mathbb{C}^N$, consider
\begin{align}
    \vec{x}^\dagger \mM \vec{x} &= \sum_{ij} x_i^* x_j \mM_{ij} \nonumber \\
        &= \langle \left[ \sum_i x_i^* A_i^\dagger, \left[ H, \sum_j x_j A_j \right] \right]  \rangle \nonumber \\
        &= \langle [B^\dagger, [H, B]] \rangle\,,\quad B \coloneqq  \sum_j x_j A_j\,.
\end{align}
Expanding the thermal expectation value $\langle O \rangle = \frac{1}{Z} \Trace [\exp(-\beta H_\text{MF}) O]$, where $\beta$ is the inverse temperature and $Z$ the partition function, and the commutators, we obtain

@@

@@

>>>>

In this section, we will be investigating four different diagonal Green's functions $G_{AA^\dagger}(\omega + i0^+)$, namely with
\begin{subequations}
    \label{eqn:resolvent_bases}
    \begin{align}
        A_\text{Higgs} &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( f_{\vk} + f_{\vk}^\dagger \right)  
            = \int \dgamma \left( f_{\gamma} + f_{\gamma}^\dagger \right) \,,\\
        A_\text{Phase} &= \frac{i}{\sqrt{N}} \sum_{\vk} \left( f_{\vk} - f_{\vk}^\dagger \right) = \int \dgamma \left( f_{\gamma} - f_{\gamma}^\dagger \right) \,,\\
        A_\text{CDW}   &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( g_{\vk \up} + g_{\vk \down} \right) = \int \dgamma \left( g_{\gamma \up} + g_{\gamma \down} \right) \,,\\
        A_\text{AFM}   &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( g_{\vk \up} - g_{\vk \down} \right) = \int \dgamma \left( g_{\gamma \up} - g_{\gamma \down} \right) \,,
    \end{align}
\end{subequations}
where $N$ is the number of lattice sites in the system. 
Each one of these operators generates a different kind of collective mode.
The first operator captures the amplitude mode of the $s$-wave superconducting state, while the second one catches the phase mode \cite{Fan22}.
The remaining two capture the collective behaviour of the CDW and AFM states, respectively.

Note, that all of the aforementioned operators are Hermitian. 
This means that all spectral functions need to be completely antisymmetrical, see \autoref{sec:antisymmetry_spectral}.
For this reason, we will only show the $\omega \geq 0$ in our plots. 
Additionally, we will add a small imaginary part of $10^{-5}t$ to $\omega$ in order to visualize the occurring $\delta$ peaks.

\subsection{Classification of the spectral functions}

To begin, let us describe the the spectral function with respect to the operators above in the various phases.
A plot of those in the SC and CDW phase, close to the phase transition, is shown in \autoref{fig:resolvent_overview_SC}.
We chose $U=-2.5t$ and $V=\pm0.1t$ for both lattices.

@@

To further classify the aforementioned peaks that lie outside the two-particle continuum,
we turn towards the real part of the Green's function.
We plot it in close proximity to the peaks with a double logarithmic scale and fit the result linearly $y = ax + b$ to obtain its power-law behaviour.
The real part is related to the imaginary part via the Kramers-Kronig relations.
Specifically, if the imaginary part is a $\delta$ distribution, the real part will be proportional to $1/\omega$, i.e., $a=-1$ and the peak has the weight $w_0 = \exp(b)$.
Furthermore, an exponent of $a=-2$ corresponds to the derivative of the $\delta$ distribution, i.e., $\delta'(\omega - \omega_0)$.

The plots for the phase peak and the CDW peak in the SC phase can be found in \autoref{sec:fit_greens_functions}.
The SC peak in the CDW phase behaves similarly to the latter and is therefore not shown explicitly.
We get a good agreement with the power-law behaviour near the peaks with the exponents $a=-2$ (phase) and $a=-1$ (CDW), respectively.
Therefore, we conclude that the CDW peak corresponds to a $\delta$ distribution, while the phase peak is its derivative.
Note, that the latter is located at $\omega=0$ of a perfectly asymmetrical bosonic spectral function and therefore must have 0 weight.
This is given here, because $\int \delta'(\omega) \mathrm{d}\omega = 0$.
The peak itself can be interpreted as two $\delta$ distributions at $0^\pm$.

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/resolvent_overview_AFM_CDW.pdf}
    \caption{Spectral functions with respect to the operators listed in \eqref{eqn:resolvent_bases}.
    The left column shows the results for the square lattice while the right column shows the results for the simple cubic lattice.
    The shaded area marks the two-particle continuum.
    The panels (a) and (b) show the spectral functions in the CDW and AFM phases, respectively.}
    \label{fig:resolvent_overview_AFM}
\end{figure*}

Next, we will investigate the Higgs peak, this time around, turning towards the spectral function itself. 
We see in \autoref{fig:resolvent_overview_SC} that the Higgs and phase spectral functions become indistinguishable towards the upper half of the two-particle continuum.
Therefore, we interpret the latter as a background to the Higgs peak itself and compute the difference between both spectral functions.
Using the same methodology as before we obtain that the exponent is $a\approx-0.5$, indicating that the peak behaves as $1 / \sqrt{\omega - \omega_0}$.

Lastly, we investigate the same spectral functions close to the AFM-CDW phase transition, depicted in \autoref{fig:resolvent_overview_AFM}.
We choose $V=1t$ for both lattices and $U=zV \pm 0.1t$, where $z$ is the coordination number.
Again, the both SC-related spectral functions are identical in both the CDW and AFM phase.
Within the continuum their weight is shifted towards the upper edge and on the square lattice only, there is a peak above the continuum.
In the AFM phase, both the CDW and AFM spectral functions have a peak close, but yet below, the two-particle continuum.
The AFM peak is at a lower energy than the CDW one.
In the CDW phase, this behaviour is reversed, i.e., the CDW peak lies lower.
Using the same analysis as before, we conclude that each of these peaks is a $\delta$-peak.

@@

TODO

@@

We show that $\mathcal{M}$ is positive semidefinite if the system is in thermal equilibrium.
For any $\vec{x} \in \mathbb{C}^N$, consider
\begin{align}
    \vec{x}^\dagger \mM \vec{x} &= \sum_{ij} x_i^* x_j \mM_{ij} \nonumber \\
        &= \langle \left[ \sum_i x_i^* A_i^\dagger, \left[ H, \sum_j x_j A_j \right] \right]  \rangle \nonumber \\
        &= \langle [B^\dagger, [H, B]] \rangle\,,\quad B \coloneqq  \sum_j x_j A_j\,.
\end{align}
For $\mM$ to be positive semidefinite the expression above must not be negative.
Therefore, we need to show, that any diagonal element of $\mM$ is nonnegative.
To do this, we expand the commutators and the thermal expectation value $\langle O \rangle = \frac{1}{Z} \Trace [\exp(-\beta H_\text{MF}) O]$, 
where $\beta$ is the inverse temperature and $Z$ the partition function.
This procedure yields

@@

This concludes our proof. Note, that the only assumption made, is that the expectation is to be taken with respect to the thermal equilibrium.

@@

\section{Fits to the Green's functions}
\label{sec:fit_greens_functions}
